---
title: "Get species information"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get species information}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```



## Calculating rmax

The demographic potential of a species is limited. The intrinsic rate of 
increase $rmax$ is the maximum increase in log population size that a species 
can attain in a year. According to Sinclair (2003), it is related to the body 
mass of adult females $W$  by the formula:


$$ 1.375*W^{-0.315} $$


Body masses are found in the literature in publications such as Kingdon & 
Hoffman (2013), Cornelis _et al._ (2014), Illius & Gordon (1992), 
Sinclair (1996), Suraud _et al._ (2012), or Foley & Faust (2010). 
Alternatively, $rmax$ can be obtained from specific demographic analyses. In the 
following table, we have listed the $rmax$ values obtained with one or the other 
method giving precedence to specific analyses when available.  



```{r 'rmax'}
species_name <- c("Impala", "Tiang", "Blue wildebeest", "Roan", "Buffalo", 
                  "Eland", "Giraffe", "Elephant")

adult_female_body_mass <- c(55, 127, 230, 250, 400, 450, 702, 2873)

species <- data.frame(adult_female_body_mass, row.names = species_name)

species$rmax <- 1.375 * adult_female_body_mass ^ (-0.315)
species["Eland", "rmax"]    <- 0.150                 # Sinclair (1996)
species["Elephant", "rmax"] <- 0.07                  # Foley & Faust (2010)
species["Giraffe", "rmax"]  <- 0.125                 # Suraud et al. (2012)

species
```



## Building conversion table

Aerial and ground counts are not directly comparable because some species are 
better detected from the ground and others from the air. It is generally 
considered that small light species are better detected from the ground while 
large dark species are better detected from the air. We took advantage of series 
of counts carried out in parallel using the two field methods to calculate 
conversion factors to be applied to aerial counts to obtain ground count 
equivalents. This permits reconciling the two types of counts within a mixed 
series. We also specify the preferred field method for each category of species.



```{r 'conversion_fact'}
categories <- c("Medium light and brown species (20-150kg)",
                "Large light and brown species (>150kg)",
                "Large dark (>150kg)", "Giraffe", "Elephant")

short_names        <- c("MLB", "LLB", "LD", "Giraffe", "Elephant")
conversion_facts   <- c(6.747, 2.302, 0.561, 3.011, 0.659)
pref_field_methods <- c("G", "G", "A", "A", "A")

category_info <- data.frame("category"          = categories,
                            "acronym"           = short_names,
                            "conversion_fact"   = conversion_facts, 
                            "pref_field_method" = pref_field_methods)
category_info
```



## Categorizing species

Here we relate the species to a color/size category.


```{r}
species$"category" <- c("MLB", "MLB", "LLB", "LLB", "LD", "LLB", "Giraffe", 
                        "Elephant")
species
```
